### 2021-10-12 tikz/xlrm.tex

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}
\usepackage[ttscale=0.8]{libertine}
\usepackage{physics}

\usepackage{xcolor}
\definecolor{TurquoiseGreen}{HTML}{B8D5B8}
\definecolor{TealBlue}{HTML}{598392}
\definecolor{YellowOrange}{HTML}{FFB140}
\definecolor{RichBlack}{HTML}{011627}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{uncertainty} = [rectangle, draw, fill=red!30, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{lever} = [rectangle, draw, fill=yellow!20, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{relation} = [rectangle, draw, fill=gray!20, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{metric} = [rectangle, draw, fill=blue!20, text=RichBlack, text width=6.5em, text centered]

\tikzstyle{line} = [draw, very thick, color=RichBlack, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    \node [relation] (relationships) at (0, 0) {System\\Relationships\\(R)};
    \node [uncertainty] (surge) at (-3.5, 2.5) {Storm Surge\\{\footnotesize ($J$ draws from 1 posterior)}};
    \path [line] (surge) to (relationships);
    \node [uncertainty] (slr) at (-3.5, -1) {Sea Level Rise\\{\footnotesize ($J$ Simulations)}};
    \path [line] (slr) to (relationships);
    \node [lever] (elevate) at (0, 2.5) {How high to elevate?};
    \path [line] (elevate) to (relationships);
    \node [metric] (ccost) at (3.5, 1.25) {Construction Costs};
    \path [line] (relationships) to (ccost);
    \node [metric] (fcost) at (3.5, -1.25) {Flood Damages};
    \path [line] (relationships) to (fcost);

    \node[uncertainty] (rcp) at (-4.5, -2.5) {RCP Scenario\\{\footnotesize (4 Scenarios)}};
    \node[uncertainty] (ice) at (-4.5, 0.75) {Ice Sheet Dynamics\\{\footnotesize (3 Models)}};
    \path [line] (rcp) to (slr);
    \path [line] (ice) to (slr);

\end{tikzpicture}

\end{document}