### 2021-10-28 tikz/fema-approach.tex
a major reorganization is underway, this is just an intermediate saved point

\documentclass[tikz, border=1pt]{standalone}

\usepackage{libertine}

\usepackage{tikz}
\usetikzlibrary{
  shapes,
  arrows,
  shapes.geometric,
  positioning,
}

% Define block styles
\tikzstyle{model} = [
  rectangle,
  align=left,
]
\tikzstyle{arrow} = [draw, -latex']

% set consistent font size
\tikzset{
    every picture/.style={font issue=\small},
    font issue/.style={execute at begin picture={#1\selectfont}}
 }

\begin{document}
\begin{tikzpicture}[auto, family=\sffamily]
  \node[model] (a) at (0,0) {
    (a)\\
    \includegraphics[height=2.5in]{../plots/historic_surge.pdf}
  };
  \node[model] (b) at (8.5,0) {
    (b)\\
    \includegraphics[height=2.5in]{../plots/surge_cdf.pdf}
  };
\end{tikzpicture}
\end{document}