### 2021-12-10 papers/agu21/combining.tex

\begin{block}{synthesizing pdfs for decision relevance}
    If there are only a few PDFs, then qualitative comparisions (\eg, \cref{fig:tradeoffs}) may be sufficient.
    With \emph{many} PDFs (\eg, \cref{fig:boxplots}), further synthesis is needed.

    We consider the \emph{common case} of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{bayes-rdm.pdf}
        \caption{
            Sketch of the ``Prior Model Averaging'' framework.
        }\label{fig:flowchart}
    \end{figure}
\end{block}