### 2021-12-10 papers/agu21/cut-paste.tex

Flood elevation is a widely used strategy for managing household flood risk.
Yet official guidance from \gls{fema}, \textbf{and regulated through local building codes} in many areas, uses a ``one size fits all'' approach that is silent on important questions.
\begin{framed}
    \begin{description}
        \item[Q1] How to adapt recommendations to a particular house's structure and a particular household's objectives?
        \item[Q2] How to incorporate nonstationary flood risk into decision support and building codes?
        \item[Q3] Which scenarios and model structures should we design for?
    \end{description}
\end{framed}
\textbf{These issues challenge efforts to incorporate climate change into engineering codes more generally.}