### 2021-12-10 papers/agu21/not-equally-plausible.tex

\begin{block}{we're always }
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{lsl_priors.pdf}
        \caption{
            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution.
            This may or may not accurately reflect available information.
        }
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{inference_weights.pdf}
        \caption{
            The prior model averaging approach (\cref{fig:flowchart}) re-weights PDFs based on their distance, and that of other PDFs available, from the (ideally coproduced!) prior.
        }
    \end{figure}
\end{block}