### 2021-12-10 papers/agu21/scenario-matters.tex

\begin{block}{it matters which pdf you choose!}
    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes (see ref.~\cite{gelman_philosophy:2013} for a more philosophical discussion).
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byscenario.pdf}%
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_iqr_byscenario.pdf}
            \caption{
                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the model selected.
            }
            \label{fig:tradeoffs}
        \end{figure}
    \end{framed}
    Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement
\end{block}