### 2021-12-10 tikz/bayes-rdm.tex

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}

\usepackage{physics}
\usepackage{libertine}

\usepackage{xcolor}
% FORZA GRANDE ROMA!!! https://teamcolorcodes.com/a-s-roma-color-codes/
\definecolor{giallo}{HTML}{F0BC42}
\definecolor{rosso}{HTML}{8E1F2F}
\definecolor{grigio}{HTML}{CACACC}
\definecolor{nero}{HTML}{000000}
\definecolor{thirds}{HTML}{31467A}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, minimum width=5.5em, text centered, align=center]
\tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=center]
\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=rosso, text=grigio, minimum width=6em, text centered,  minimum height=4em, align=center]
\tikzstyle{check} = [text=thirds, align=center]

\tikzstyle{line} = [draw, very thick, color=nero, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    \node [block] (decisions) at (0, 3) {
        Candidate decisions\\
        $\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$
    };
    \node [block] at (-4, 3) (models) {
        Multiple PDFs\\
        $\qty{M_1, \ldots, M_K} \in \mathcal{M}$
    };
    \node [block] at (-4, 0) (simulations) {
        Simulated SOWs\\
        $\qty{s_1, \ldots, s_{N_k}}_k$
    };
    \path [line] (models) -- (simulations);
    \node [function] (model) at (0, 0) {
        System model\\
        $u_{ij} = f(x_i, s_j)$
    };
    \path [line] (simulations) -- (model);
    \path [line] (decisions) -- (model);

    \node [derived] (combined) at (4, -3) {
        Conditional distributions\\
        over outcomes\\
        $\hat{p} \qty(u | x_i, M_k)$
    };
    \path [line] (model) -- (combined);

    \node [function] (prior) at (-4, -3) {
        Prior\\
        over SOWs\\
        $p\qty(s)$
    };
    \node [derived] (weights) at (0, -3) {
        Model weights\\
        $w_1, \ldots, w_K$
    };
    \path [line] (simulations) -- (weights);
    \path [line] (prior) -- (weights);

    \node [derived] (aggregated) at (4, -6) {
        Aggregated distribution\\
        over outcomes\\
        $\hat{p}(u | x_i)$
    };
    \path [line] (combined) -- (aggregated);
    \path [line] (weights) -- (aggregated);
\end{tikzpicture}

\end{document}