### 2021-12-10 papers/agu21/not-equally-plausible.tex

\begin{block}{different beliefs emphasizes different possibilities}
    The prior model averaging approach (\cref{fig:flowchart}) assigns a weight to each PDF based on their
    \begin{itemize}
        \item The prior can't be ``right''  \cite{gelman_workflow:2020,gelman_philosophy:2013} or ``scenario neutral'' \cite{quinn_exploratory:2020}
        \item let's at least \textbf{make assumptions transparent}!
    \end{itemize}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{lsl_priors.pdf}
        \caption{
            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution; this may or may not reflect available information.        }
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{inference_weights.pdf}
        \caption{
            Weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
            Horizontal line shows the weight given to each under an implicit uniform assumption.
        }
    \end{figure}
\end{block}