### 2021-12-10 papers/agu21/house-elevation.tex

\begin{block}{house elevation}
    Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \cite{zarekarizi_suboptimal:2020,xian_elevation:2017}.
    \vspace*{-1em}
    \begin{description}
        \item[Q1] How to adapt guidance to building characteristics or household preferences?
        \item[Q2] How does nonstationary hazard change guidance?
        \item[Q3] Which model of nonstationary hazard to use?
    \end{description}
    \begin{framed}
        \begin{figure}
            \centering
            \subcaptionbox{Selinsgrove, PA}{\includegraphics[height=3.35in]{img_selinsgrove.jpeg}}%
            \hfill%
            \subcaptionbox{Houston, TX}{\includegraphics[height=3.35in]{img_braeswood.png}}\\
            \subcaptionbox{Plaquemines Parish, LA}{\includegraphics[height=3.3in]{epstein-2020-la.jpeg}}%
            \hfill%
            \subcaptionbox{Dare County, NC}{\includegraphics[height=3.3in]{nichols-ocracoke.jpeg}}%
            \caption{
                \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols / PSU.
            }
            \label{fig:images}
        \end{figure}
    \end{framed}
    These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
\end{block}