### 2021-12-10 papers/agu21/multiple-pdf.tex

\begin{block}{the multiple pdf problem}
    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
    Including additional models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{msl_boxplots.pdf}
            \caption{
                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017,ruckert_coastal:2019}.
                We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
            }\label{fig:boxplots}
        \end{figure}
    \end{framed}
\end{block}