### 2022-03-16 tikz/xlrm.tex
edits esp base SLR priors on NOAA report

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}
\usepackage[ttscale=0.8]{libertine}
\usepackage{physics}

\usepackage{xcolor}
\definecolor{giallo}{HTML}{F0BC42}
\definecolor{rosso}{HTML}{8E1F2F}
\definecolor{grigio}{HTML}{CACACC}
\definecolor{nero}{HTML}{000000}
\definecolor{thirds}{HTML}{31467A}
\definecolor{orange}{HTML}{F18101}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{uncertainty} = [rectangle, draw, fill=rosso, text=grigio, text width=6.5em, text centered]
\tikzstyle{lever} = [rectangle, draw, fill=giallo, text=nero, text width=6.5em, text centered]
\tikzstyle{relation} = [rectangle, draw, fill=thirds, text=grigio, text width=6.5em, text centered]
\tikzstyle{metric} = [rectangle, draw, fill=orange, text=nero, text width=6.5em, text centered]

\tikzstyle{line} = [draw, very thick, color=nero, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    \node [font=\color{rosso}\selectfont] at (-6, 2.5) {\Large Uncertainties};
    \node [relation] (relationships) at (0, -0.5) {Engineering and economic models};
    \node [uncertainty] (surge) at (-3, 1.5) {Storm surge};
    \path [line] (surge) to (relationships);

    \node [font=\color{thirds}\selectfont] at (0, -2) {\Large Relationships};
    \node [uncertainty] (slr) at (-3.5, -0.5) {Sea level};
    \path [line] (slr) to (relationships);

    \node [font=\color{giallo}\selectfont] at (0, 2.5) {\Large Levers};
    \node [lever] (elevate) at (0, 1.5) {How high to elevate?};
    \path [line] (elevate) to (relationships);

    \node [font=\color{giallo}\selectfont] at (3.5, 2.5) {\Large Metrics};
    \node [metric] (ccost) at (3.5, 1) {Up-front\\(construction)\\cost};
    \path [line] (relationships) to (ccost);
    \node [metric] (fcost) at (3.5, -1) {Lifetime expected damages};
    \path [line] (relationships) to (fcost);

    \node[uncertainty] (rcp) at (-6, -1.5) {RCP scenario\\{\footnotesize (4 Scenarios)}};
    \node[uncertainty] (ice) at (-6, 0.75) {Ice sheet dynamics\\{\footnotesize (4 Models)}};
    \path [line] (rcp) to (slr);
    \path [line] (ice) to (slr);

\end{tikzpicture}

\end{document}