### 2022-03-30 tikz/bayes-rdm.tex

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}

\usepackage{physics}
\usepackage{libertine}

\usepackage{xcolor}
% FORZA GRANDE ROMA!!! https://teamcolorcodes.com/a-s-roma-color-codes/
\definecolor{giallo}{HTML}{F0BC42}
\definecolor{rosso}{HTML}{8E1F2F}
\definecolor{grigio}{HTML}{CACACC}
\definecolor{nero}{HTML}{000000}
\definecolor{thirds}{HTML}{31467A}
\definecolor{orange}{HTML}{F18101}
\definecolor{deepblue}{HTML}{CA3CFF}

% https://tex.stackexchange.com/questions/200838/how-do-i-make-dots-more-compact-or-are-there-alternatives
\newcommand{\myDots}{\ifmmode\mathinner{\ldotp\kern-0.2em\ldotp\kern-0.2em\ldotp}\else.\kern-0.13em.\kern-0.13em.\fi}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, minimum width=5.5em, text centered, align=center]
\tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=left]
\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=rosso, text=grigio, minimum width=6em, text centered,  minimum height=4em, align=left]

\tikzstyle{line} = [draw, very thick, color=nero, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    % label the various sections
    \path [fill=orange!30] (-6.75, 4.8) -- (5.75, 4.8) -- (5.75, -4) -- (-6.75, -4) -- cycle;
    \node [color=thirds] (s5) at (-4, -3) {\textbf{III. Synthesize}};

    \path [fill=rosso!30] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- cycle;
    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{II. Condition}};

    \path [fill=thirds!30] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- cycle;
    \node [color=thirds] (s3) at (0, 4) {\textbf{I. Explore}};

    \node [block, label={[]225:{\footnotesize (d)}}] at (-4, 3) (models) {
            Multiple models\\
            $\vb{M} = \qty{M_1, \ldots, M_K}$\\
            for $p(\Omega | M_k \in \mathcal{M})$
        };
    \node [block, label={[]225:{\footnotesize (a)}}] (decisions) at (0, 3) {
            Candidate decisions\\
            $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$
        };
    \node [block, label={[]225:{\footnotesize (b)}}] at (-4, 0) (simulations) {
            Many possible futures\\
            $\vb{s}=\qty{s_1, \ldots, s_J}$\\
            where $s_j \in \Omega$
        };
    \path [line] (models) -- (simulations);
    \node [function, label={[]45:{\footnotesize (c)}}] (model) at (0, 0) {
            System model\\
            $u_{ij} = f(x_i, s_j)$
        };
    \path [line] (simulations) -- (model);
    \path [line] (decisions) -- (model);

    \node [derived, label={[]0:{\footnotesize (f)}}] (weights) at (0, -3) {
            Weights\\
            $\vb{w} = \qty{w_1, \ldots, w_J}$\\
            where $w_j \in (0, 1)$
        };
    \path [line] (simulations) -- (weights);

    \node [derived, label={[]270:{\footnotesize (e)}}] (combined) at (4, 0) {
            Conditional\\
            distribution\\
            over outcomes\\
            $p(u | x_i, M_k)$
        };
    \path [line] (model) -- (combined);
    \path [line] (weights) -- (combined);

\end{tikzpicture}

\end{document}